---
title: "Juvenile development time"
output: 
  workflowr::wflow_html:
    code_folding: hide 
editor_options:
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(warning=FALSE, message=FALSE)
```

## Load packages
```{r}
library(tidyverse)
library(ggridges)

library(coxme)
library(brms)
library(tidybayes)

library(kableExtra)
library(knitrhooks) # install with devtools::install_github("nathaneastwood/knitrhooks")
library(showtext)

output_max_height() # a knitrhook option

options(stringsAsFactors = FALSE)

# set up nice font for figure
nice_font <- "Lora"
font_add_google(name = nice_font, family = nice_font, regular.wt = 400, bold.wt = 700)
showtext_auto()

```

## Load data
```{r, results='hide'}
# load eclosion data
eclosion_wide <- read.csv("data/1.eclosion_wide.csv")

# convert data to 'long' format
eclosion_long <- reshape2::melt(eclosion_wide, id.vars = "DAY")

# add ID columns
eclosion_long$YEAR <- factor("2015")
eclosion_long$variable <- paste(eclosion_long$variable, eclosion_long$YEAR, sep = "_")
eclosion_long$ID <- gsub(pattern = "_m|f_", replacement = "", x = eclosion_long$variable)

# calculate days to eclosion starting from seeding day
eclosion_long <- eclosion_long %>% 
  separate(variable, c("LINE", "SEED", "VIAL", "SEX", "YEAR")) %>% 
  mutate(SEED_DAY = case_when(SEED == 'A' ~ "2015-7-20", 
                              SEED == 'B' ~ "2015-7-21",
                              SEED == 'C' ~ "2015-7-22"),
         START_DAY = case_when(SEED == 'A' ~ "2015-8-01", 
                               SEED == 'B' ~ "2015-8-02",
                               SEED == "C" & 
                                 LINE == "M1"| LINE == "M2"| LINE == "M3"| LINE == "M4" ~ "2015-8-03",
                               TRUE ~ "2015-8-04"),
         DAY_ZERO = as.numeric(as.Date(START_DAY) - as.Date(SEED_DAY)),
         DAY = DAY + DAY_ZERO)


# remove rows where no flies eclosed
eclosion_long <- subset(eclosion_long, value>0)

# add event 
eclosion_long$EVENT <- 1

# calulate number not eclosed for each vial (out of 100)
uneclosed <- as.data.frame(eclosion_long %>% 
                             group_by(ID, LINE, SEED, VIAL) %>% 
                             summarise(SEX = "fm",
                                       DAY = 21,
                                       eclosing = 100 - sum(value),
                                       EVENT = 0,
                                       YEAR = "2015"))


# calculate proportion eclosed of each sex to give weights to uneclosed
p <- eclosion_long %>% group_by(SEX) %>% summarise(S = sum(value))
prop.male <- p$S[2]/(p$S[2]+p$S[1])
prop.female <- p$S[1]/(p$S[2]+p$S[1])

# assign sex to uneclosed in each vial based on those that have eclosed... i.e. slighly more females already eclosed
uneclosed.male <- mutate(uneclosed, 
                            SEX = "m", eclosing = round(uneclosed$eclosing*prop.female))

uneclosed.female <- mutate(uneclosed, 
                              SEX = "f", eclosing = round(uneclosed$eclosing*prop.male))

colnames(eclosion_long)[7] <- "eclosing"

# reorder columns
eclosion_long <- eclosion_long %>% select(ID, DAY, SEX, LINE, SEED, VIAL, YEAR, eclosing, EVENT)

# bind eclosed data to calculated uneclosed
eclosion_long <- rbind(eclosion_long, uneclosed.male, uneclosed.female)
eclosion_long <- eclosion_long[order(eclosion_long$ID, eclosion_long$SEX), ]

# add treatment variable 
eclosion_long$TRT <- factor(substr(eclosion_long$ID, 1, 1))

# remove vials seeded with more than 100 larvae
#unique(eclosion.dat[which(eclosion.dat$eclosing < 0), "ID"]) # 4 vials overseeded
eclosion.dat.trim <- eclosion_long %>% 
  filter(ID %in% eclosion_long[which(eclosion_long$eclosing < 0), "ID"] == FALSE)

# expand data frame so each row is a single fly
ecl.dat <- reshape::untable(eclosion.dat.trim[ ,c(1:7, 9, 10)], 
                           num = eclosion.dat.trim[, 8])

# load wing length data
wing_length <- read.csv("data/1.wing_length.csv") %>% 
  filter(Side == 'L') %>% 
  # scale wing vein length to make effect size comparisons with other data sets?
  mutate(Length = as.numeric(scale(Length)))

# add replicate
wing_length$LINE <- paste0(wing_length$Treatment, substr(wing_length$Rep, 2, 2))

```


## Inspecting the raw data

```{r, fig.height=3}
ecl.dat %>% 
  filter(EVENT == 1) %>% 
  select(TRT, SEX, DAY) %>% 
  mutate(TRT_SEX = paste0(TRT, SEX),
         SEX = factor(ifelse(SEX == "m", "Males", "Females"))) %>% 
  ggplot(aes(x = DAY, y = SEX, fill = TRT_SEX)) +
  geom_boxplot() +
  scale_fill_manual(values = c("pink", "skyblue", "red", "blue"), name = "",
                    labels = c('Monogamy Females', 'Monogamy Males',
                               'Polandry Females', 'Polandry Males')) +
  labs(x = 'Eclosion time (days)', y = 'Sex') +
  theme_bw() +
  NULL

```

# Survival analysis 
We modelled juvenile development time using survival analysis. We measured the time in days from 1st instar larvae until eclosion (`EVENT` = 1) upon which flies were stored in ethanol before counting. Of the initially seeded 100 flies per vial, the remaining flies not emerging after two consecutive days of no observed eclosions were right censored (`EVENT` = 0) on the last observation day. In total 14400 larvae were seeded (100 larvae x 2 `Treatment` x 4 `LINE` per `Treatment` x 6 `VIAL` per `LINE` x 3 `SEED` days). For P1 only three vials were seeded on day B so we seeded 3 additional vials on day C. Four vials were seeded with too many larvae and excluded from analysis. In total `r ecl.dat %>% count(EVENT) %>% filter(EVENT == 1) %>% pull(n)` flies eclosed during the observation period leaving `r ecl.dat %>% count(EVENT) %>% filter(EVENT == 0) %>% pull(n)` individuals to be right censored on day 9. 

### Censoring
Censored flies were assigned sex based on the observed sex ratio of eclosees assuming an equal (50:50) sex ratio of larvae seeded to each vial at the beginning of the experiment. We calculate the number of number of males and females that did eclose and subsequently assign sex to the remaining (uneclosed) individuals of unknown sex based on the proportion of individuals of each sex that did emerge. For example, if 70 flies were counted eclosing from a vial with 40 females and 30 males, we then designate the remaining 30 flies as 10 females and 20 males and so on so that each vial ends with 50 females and 50 males some of which are right censored (`EVENT` = 0).


### Kaplan-Meier survival curve
First we plot Kaplan-Meier survival curves without considering our full experimental design. 

```{r, fig.width=4.5, fig.height=3.8}
survminer::ggsurvplot(survfit(Surv(DAY, EVENT) ~ TRT + SEX, data = ecl.dat),
                      conf.int = TRUE,
                      risk.table = FALSE,
                      linetype = "SEX",
                      palette = c("pink", "skyblue", "red", "blue"),
                      fun = "event",
                      xlim = c(12, 21),
                      xlab = "Days",
                      legend = 'right',
                      legend.title = "",
                      legend.labs = c("M \u2640","M \u2642",'E \u2640','E \u2642'),
                      break.time.by = 2,
                      ggtheme = theme_bw())
#ggsave(filename = 'figures/eclosion.pdf', width = 5.5, height = 5, dpi = 600, useDingbats = FALSE)

```
**Figure X:** Kaplan-Meier curve for eclosion time (in days) for flies in each treatment and sex. +'s indicate censored individuals (n = `r ecl.dat %>% count(EVENT) %>% filter(EVENT == 0) %>% pull(n)`).

### Median eclosion times

```{r}
summary(survfit(Surv(DAY, EVENT) ~ TRT + SEX, data = ecl.dat))$table %>% 
  as_tibble() %>% 
  mutate(Treatment = c('M', 'M', 'E', 'E'),
         Sex = c('Female', 'Male', 'Female', 'Male')) %>% 
  mutate(`Median (± 95% CI)` = paste0(median, ' (', `0.95LCL`, '-', `0.95UCL`, ')')) %>% 
  dplyr::select(Treatment, Sex, N = records, `N events` = events, `Median (± 95% CI)`) %>%
  kable() %>% 
  kable_styling(full_width = FALSE)

```

### Check proportional hazards assumption
Next we need to check that the proportional hazards assumption is not violated before fitting the model, where crossing hazards (lines) indicate violation of the proportional hazards assumption. 
```{r, fig.width=4.5, fig.height=3.8}
survminer::ggsurvplot(survfit(Surv(DAY, EVENT) ~ TRT + SEX, data = ecl.dat),
                      conf.int = TRUE,
                      risk.table = FALSE,
                      linetype = "SEX",
                      palette = c("pink", "skyblue", "red", "blue"),
                      fun = "cloglog",
                      xlim = c(13, 21),
                      legend = 'right',
                      legend.title = "",
                      legend.labs = c("M \u2640","M \u2642",'E \u2640','E \u2642'),
                      break.time.by = 2,
                      ggtheme = theme_bw())

```
**Figure X:** ln(-ln(survival))

## Fit the model in `brms`
We fit a Cox Proportional hazards model in `brms` using `family = cox()`, with time (days) to event (eclosion) as the response and sexual selection treatment (`TRT`; Monogamy or Elevated polyandry), `SEX` (female or male) and their interaction as predictors with Seed day as a covariate. [See here](https://bookdown.org/content/4857/god-spiked-the-integers.html) for a helpful explanation on fitting survival models in `brms`. We also include replicate treatment as a random intercept term for each of the 8 lines and a random slope term to allow the effect of selection treatment to vary across replicate lines. We also include vial `ID` as a random intercept term as individuals emerging from the same vial may show a correlated response. 

#### Define priors
We set conservative normal priors on the fixed effects (mean = 0, sd = 1) and half Cauchy priors on the random effects - `LINE` and vial `ID` - (mean = 0, scale = 0.1). All other priors were left at the default in `brms`.

#### Run the model
The model is run over 4 chains with 5000 iterations each (with the first 2500 discarded as burn-in), for a total of 2500*4 = 10,000 posterior samples. Note that some of the `brms` functionality is not currently available for models using the `cox` family (e.g. posterior predictive checks).
```{r}

if(!file.exists("output/cox_brms.rds")){
  
  cox_brm <- brm(DAY | cens(1 - EVENT) ~ TRT * SEX + SEED + (TRT|LINE) + (1|ID),
               iter = 5000, chains = 4, cores = 4,
               prior = c(set_prior("normal(0,1)", class = "b"),
                         set_prior("cauchy(0,0.1)", class = "sd")),
               control = list(max_treedepth = 20,
                              adapt_delta = 0.999),
               data = ecl.dat, family = cox())
  
  #saveRDS(cox_brm, "output/cox_brms_noslope.rds") # save with no random slope term
  saveRDS(cox_brm, "output/cox_brms.rds")
} else {
  cox_brm <- readRDS('output/cox_brms_noslope.rds')
}

```

### Table of model parameter estimates - eclosion time {.tabset}

#### Formatted table 
This tables shows the fixed effects estimates on eclosion time. The p column shows 1 - minus the “probability of direction”, i.e. the posterior probability that the reported sign of the estimate is correct given the data and the prior; subtracting this value from one gives a Bayesian equivalent of a one-sided p-value. Click the next tab to see a complete summary of the model and its output.
```{r}
hyp_test <- bind_rows(
  hypothesis(cox_brm, 'TRTP = 0')$hypothesis,
  hypothesis(cox_brm, 'SEXm = 0')$hypothesis,
  hypothesis(cox_brm, 'TRTP:SEXm = 0')$hypothesis,
  hypothesis(cox_brm, 'SEEDB = 0')$hypothesis,
  hypothesis(cox_brm, 'SEEDC = 0')$hypothesis
) %>% 
  mutate(Parameter = c('Treatment (E)', 'Sex (M)', 'Treatment (E) x Sex (M)', 'Seed (B)', 'Seed (C)'),
         across(2:5, round, 3)) %>% 
  relocate(Parameter, Estimate, Est.Error, CI.Lower, CI.Upper, Star)

pvals <- bayestestR::p_direction(cox_brm) %>% 
  as.data.frame() %>%
  mutate(vars = map_chr(str_split(Parameter, "_"), ~ .x[2]),
         p_val = 1 - pd, 
         star = ifelse(p_val < 0.05, "\\*", "")) %>%
  select(vars, p_val, star)

hyp_test %>% 
  mutate(vars = c('TRTP', 'SEXm', 'TRTP:SEXm', 'SEEDB', 'SEEDC')) %>% 
  left_join(pvals %>% filter(vars != 'Intercept'), 
            by = c("vars")) %>% 
  select(Parameter, Estimate, Est.Error, CI.Lower, CI.Upper, `p` = p_val, star) %>% 
  rename(` ` = star) %>%
  mutate(p = ifelse(p > 0.001, round(p, 3), '< 0.001')) %>% 
  kable() %>% 
  kable_styling(full_width = FALSE)

```

#### Complete output from `summary.brmsfit()`
```{r}
cox_brm
```

### Posterior effect size of treatment on eclosion time for each sex
As `posterior_eprid()` is not available for `brms` models using the `cox` family, we manually calculate the estimates for each group from the posterior predictions. The $\beta$ coefficients from a Cox model measure the impact of covariates and give an estimate of the effect size ([see here](http://www.sthda.com/english/wiki/cox-proportional-hazards-model)). Taking the exponent of the coefficients give the hazard ratio. In short, hazard ratios give the probability of the event occurring compared to the 'control' group, in our case compared to Monogamy females, where:

* Hazard ratio = 1 ($\beta$ = 0): no effect
* Hazard ratio > 1 ($\beta$ > 0): reduced hazard (higher probability of eclosion)
* Hazard ratio < 1 ($\beta$ < 0): increased hazard (lower probability of eclosion)

```{r, fig.height=4, fig.width=4}

treatsex_eclosion <- posterior_samples(cox_brm) %>% 
  as_tibble() %>%
  select(starts_with("b_")) %>% 
  mutate(draw = 1:n()) %>% 
  mutate(M_f = b_Intercept,
         P_f = b_Intercept + b_TRTP,
         M_m = b_Intercept + b_SEXm,
         P_m = b_TRTP + b_SEXm + `b_TRTP:SEXm`) %>% 
  select(draw, M_f, P_f, M_m, P_m) %>% 
  pivot_longer(cols = 2:5) %>% 
  mutate(SEX = str_sub(name, -1),
         TRT = str_sub(name, 1, 1)) %>%
  select(draw, value, SEX, TRT) %>% 
  pivot_wider(names_from = TRT,
              values_from = value) %>%
  mutate(`Difference in means (Poly - Mono)` = P - M)

treatsex_eclosion %>%
  ggplot(aes(x = SEX, y = `Difference in means (Poly - Mono)`, fill = SEX)) +
  geom_hline(yintercept = 0, linetype = 2) +
  stat_halfeye() + 
  scale_fill_brewer(palette = 'Pastel1', direction = 1, name = "") +
  scale_colour_brewer(palette = 'Pastel1', direction = 1, name = "") +
  labs(y = 'Difference in means (\u03b2) between\nselection treatments (E - M)') +
  theme_bw() +
  theme(legend.position = 'none',
        strip.background = element_blank(),
        panel.grid.major.x = element_blank()) +
  NULL

```
**Figure X:** Difference in mean $\beta$ coefficients for the survival analysis on eclosion time between the selection treatments (Elevated polyandry - Monogamy).


## Posterior difference in treatment effect size between sexes {.tabset}

This section examines the treatment $\times$ sex interaction term, by calculating the difference in the effect size of the E/M treatment between sexes. We find evidence for a treatment $\times$ sex interaction, i.e. the difference in eclosion time between the sexes was greater in the E treatment than the M treatment.

### Figure
```{r, fig.height=3, fig.width=3}
treatsex_eclosion %>%
  rename(d = `Difference in means (Poly - Mono)`) %>% 
  select(draw, SEX, d) %>%
  group_by(draw) %>%
  summarise(`Difference in effect size between sexes (male - female)` = d[2] - d[1],
            .groups = "drop") %>% 
  ggplot(aes(x = `Difference in effect size between sexes (male - female)`, y = 1, fill = stat(x < 0))) +
  geom_vline(xintercept = 0, linetype = 2) +
  stat_halfeye() +
  scale_fill_brewer(palette = 'Pastel2', direction = 1, name = "") +
  theme_bw() +
  theme(legend.position = 'none',
        text = element_text(family = nice_font),
        strip.background = element_blank()) +
  ylab("Posterior density") +
  #ggsave("figures/eclosion_interaction_plot.pdf", height=4, width=6) +
  NULL
```

### Table
```{r}
treatsex_interaction_eclosion <- treatsex_eclosion %>%
  select(draw, SEX, d =  `Difference in means (Poly - Mono)`) %>%
  arrange(draw, SEX) %>%
  group_by(draw) %>%
  summarise(`Difference in effect size between sexes (male - female)` = d[2] - d[1],
            .groups = "drop") # males - females

treatsex_interaction_eclosion %>%
  rename(x = `Difference in effect size between sexes (male - female)`) %>%
  summarise(`Difference in effect size between sexes (male - female)` = median(x),
            `Lower 95% CI` = quantile(x, probs = 0.025), 
            `Upper 95% CI` = quantile(x, probs = 0.975),
            p = 1 - as.numeric(bayestestR::p_direction(x)),
            ` ` = ifelse(p < 0.05, "\\*", ""),
            .groups = "drop") %>%
  kable(digits=3) %>%
  kable_styling(full_width = FALSE)

```


```{r}
sex_eclosion <- posterior_samples(cox_brm) %>% 
  as_tibble() %>%
  select(starts_with("b_")) %>% 
  mutate(draw = 1:n()) %>% 
  mutate(M_f = b_Intercept,
         P_f = b_Intercept + b_TRTP,
         M_m = b_Intercept + b_SEXm,
         P_m = b_TRTP + b_SEXm + `b_TRTP:SEXm`) %>% 
  select(draw, M_f, P_f, M_m, P_m) %>% 
  pivot_longer(cols = 2:5) %>% 
  mutate(SEX = str_sub(name, -1),
         TRT = str_sub(name, 1, 1)) %>%
  select(draw, value, SEX, TRT) %>% 
  pivot_wider(names_from = SEX,
              values_from = value) %>%
  mutate(`Difference in means (Female - Male)` = f - m)

sex_eclosion %>%
  ggplot(aes(x = TRT, y = `Difference in means (Female - Male)`, fill = TRT)) +
  geom_hline(yintercept = 0, linetype = 2) +
  stat_halfeye() + 
  scale_fill_brewer(palette = 'Pastel1', direction = 1, name = "") +
  scale_colour_brewer(palette = 'Pastel1', direction = 1, name = "") +
  labs(y = 'Difference in means (\u03b2) between\nsexess (Female - Male)') +
  theme_bw() +
  theme(legend.position = 'none',
        strip.background = element_blank(),
        panel.grid.major.x = element_blank()) +
  NULL

```


# Body size differences (wing vein IV length) {.tabset}
We measured the length of wing vein VI as a proxy for body size to test for differences between sexes and treatments as body size may influence development time. Prior to measurement, wing images were anonymised using a custom [python script provided by Henry Barton](https://github.com/henryjuho/Drosophila_wing_length) and then decoded for statistical analysis. Wing length was scaled by subtracting the mean (across all measurements) and dividing by the standard deviation.

### Inspecting the raw data {.tabset}
#### Violin plots
```{r, fig.height=3.5, fig.width=5}
wing_length %>% 
  mutate(var = paste(Treatment, Sex)) %>% 
  ggplot(aes(x = Sex, y = Length)) +
  geom_violin(aes(fill = var), alpha = .5) + 
  geom_boxplot(aes(fill = var), width = .1, position = position_dodge(width = .9)) +
  scale_colour_manual(values = c("pink", "skyblue", "red", "blue"), name = "") +
  scale_fill_manual(values = c("pink", "skyblue", "red", "blue"), name = "",
                    labels = c('Monogamy Females', 'Monogamy Males',
                               'Polandry Females', 'Polandry Males')) +
  labs(y = 'Wing vein IV length') +
  theme_bw() +
  theme() +
  NULL

```
**Figure X:** Wing vein IV length has been scaled (subtracted the mean and divided by the standard deviation). 

#### Means and standard errors
```{r}
wing_length %>% 
  group_by(Treatment, Sex) %>% 
  summarise(Mean = mean(Length),
            `Std. Errors` = sd(Length)/sqrt(n()),
            N = n()) %>%
  mutate(Treatment = recode(Treatment, M = "Monogamy", P = 'Polyandry'),
         Sex = recode(Sex, M = "Male", F = 'Female')) %>% 
  mutate(across(2:4, round, 2)) %>% 
  kable() %>% 
  kable_styling(full_width = FALSE)
  
```

### Fitting the model for wing length in `brms`
We fit a model in `brms` to test for differences in wing length between the sexes and sexual selection treatments. We fit treatment, sex and the treatment x sex interaction as fixed effects as well as Seed day as a covariate. As above, we included replicate treatment as a random intercept for each of the 8 lines and a random slope term for selection to allow the effect of treatment to vary across replicate lines. 

#### Define priors
As above we set conservative normal priors on the fixed effects (mean = 0, sd = 1) and half Cauchy priors on the random effects - `LINE` - (mean = 0, scale = 0.1). All other priors were left at the default in `brms`.

#### Run the model
The model is run over 4 chains with 10000 iterations each (with the first 2500 discarded as burn-in), for a total of 7500*4 = 30,000 posterior samples.
```{r}

if(!file.exists("output/wing_brms.rds")){
  
  wing_brms <- brm(Length ~ Treatment * Sex + Seed + (Treatment|LINE),
                 data = wing_length,
  iter = 10000, chains = 4, cores = 4,
  prior = c(set_prior("normal(0,1)", class = "b"),
            set_prior("cauchy(0,0.1)", class = "sd")),
  control = list(max_treedepth = 20,
                 adapt_delta = 0.999)
  )
  
  #saveRDS(wing_brms, "output/wing_brms_noslope.rds") # save with no random slope term
  saveRDS(wing_brms, "output/wing_brms.rds")
} else {
  wing_brms <- readRDS('output/wing_brms_noslope.rds')
}

```

### Posterior predictive check of model fit
```{r, fig.height=2, fig.width=3}
pp_check(wing_brms)
```

### Table of model parameter estimates - body size {.tabset}

#### Formatted table 
```{r}
wing_test <- bind_rows(
  hypothesis(wing_brms, 'TreatmentP = 0')$hypothesis,
  hypothesis(wing_brms, 'SexM = 0')$hypothesis,
  hypothesis(wing_brms, 'TreatmentP:SexM = 0')$hypothesis,
  hypothesis(wing_brms, 'SeedB = 0')$hypothesis,
  hypothesis(wing_brms, 'SeedC = 0')$hypothesis,
) %>% 
  mutate(Parameter = c('Treatment (E)', 'Sex (M)', 'Treatment (E) x Sex (M)',
                       'Seed (B)', 'Seed (C)'),
         across(2:5, round, 3)) %>% 
  relocate(Parameter, Estimate, Est.Error, CI.Lower, CI.Upper, Star)

pvals <- bayestestR::p_direction(wing_brms) %>% 
  as.data.frame() %>%
  mutate(vars = map_chr(str_split(Parameter, "_"), ~ .x[2]),
         p_val = 1 - pd, 
         star = ifelse(p_val < 0.05, "\\*", "")) %>%
  select(vars, p_val, star)

wing_test %>% mutate(vars = c('TreatmentP', 'SexM', 'TreatmentP:SexM', 'SeedB', 'SeedC')) %>% 
  left_join(pvals %>% filter(vars != 'Intercept'), 
              by = c("vars")) %>% 
    select(Parameter, Estimate, Est.Error, CI.Lower, CI.Upper, `p` = p_val, star) %>% 
  mutate(p = ifelse(p > 0.001, round(p, 3), '< 0.001')) %>% 
  rename(` ` = star) %>%
  kable() %>% 
  kable_styling(full_width = FALSE)

```

#### Complete output from `summary.brmsfit()`
```{r}
wing_brms
```


### Posterior effect size of treatment on body size for each sex
We predict the mean wing vein IV length for each treatment and sex from the model averaged across the eight replicate selection lines and seeding days. The plots show the difference in posterior estimates between the E and M treatment for each sex separately. Note that females are larger than males but effect sizes are plotted for each sex separately.

```{r, fig.height=3, fig.width=3}
new <- expand_grid(Sex = c("M", "F"),
                   Treatment = c("M", "P"),
                   LINE = NA, Seed = NA) %>%
  mutate(type = 1:n())

fitted_wing <- posterior_epred(
  wing_brms, newdata = new, re_formula = NA,
  summary = FALSE, resp = 'Length') %>%
  reshape2::melt() %>% rename(draw = Var1, type = Var2) %>%
  as_tibble() %>%
  left_join(new, by = "type") %>%
  select(draw, value, Sex, Treatment)

treat_diff_wing <- fitted_wing %>%
  spread(Treatment, value) %>%
  mutate(`Difference in means (Poly - Mono)` = P - M) 

treat_diff_wing %>% 
  ggplot(aes(x = Sex, y = `Difference in means (Poly - Mono)`, fill = Sex)) +
  geom_hline(yintercept = 0, linetype = 2) +
  stat_halfeye() + 
  scale_fill_brewer(palette = 'Pastel1', direction = 1, name = "") +
  scale_colour_brewer(palette = 'Pastel1', direction = 1, name = "") +
  labs(y = 'Difference in means between\nselection treatments (E - M)') +
  theme_bw() +
  theme(legend.position = 'none',
        strip.background = element_blank(),
        panel.grid.major.x = element_blank()) +
  NULL

```
**Figure XX:** Posterior estimates of treatment effects on wing vein IV length (proxy for body size).


### Posterior difference in treatment effect size between sexes {.tabset}

This section examines the treatment $\times$ sex interaction term, by calculating the difference in the effect size of the E/M treatment between sexes. We find no evidence for a treatment $\times$ sex interaction, i.e. the treatment effects did not differ detectably between sexes.

#### Figure
```{r, fig.height=3, fig.width=3}
treat_diff_wing %>%
  rename(d = `Difference in means (Poly - Mono)`) %>% 
  select(draw, Sex, d) %>%
  group_by(draw) %>%
  summarise(`Difference in effect size between sexes (male - female)` = d[2] - d[1],
            .groups = "drop") %>% 
  ggplot(aes(x = `Difference in effect size between sexes (male - female)`, y = 1, fill = stat(x < 0))) +
  geom_vline(xintercept = 0, linetype = 2) +
  stat_halfeye() +
  scale_fill_brewer(palette = 'Pastel2', direction = 1, name = "") +
  theme_bw() +
  theme(legend.position = 'none',
        text = element_text(family = nice_font),
        strip.background = element_blank()) +
  ylab("Posterior density") +
  #ggsave("figures/wing_interaction_plot.pdf", height=4, width=4) +
  NULL
```

#### Table
```{r}
treatsex_interaction_wing <- treat_diff_wing %>%
  select(draw, Sex, d =  `Difference in means (Poly - Mono)`) %>%
  arrange(draw, Sex) %>%
  group_by(draw) %>%
  summarise(`Difference in effect size between sexes (male - female)` = d[2] - d[1],
            .groups = "drop") # males - females

treatsex_interaction_wing %>%
  rename(x = `Difference in effect size between sexes (male - female)`) %>%
  summarise(`Difference in effect size between sexes (male - female)` = median(x),
            `Lower 95% CI` = quantile(x, probs = 0.025), 
            `Upper 95% CI` = quantile(x, probs = 0.975),
            p = 1 - as.numeric(bayestestR::p_direction(x)),
            ` ` = ifelse(p < 0.05, "\\*", ""),
            .groups = "drop") %>%
  kable(digits=3) %>%
  kable_styling(full_width = FALSE)

```